---
title: "LHS 610: Week 1 HW Assignment 1"
author: "Sereen Kallerackal"
date: "27 January 2021"
output: html_notebook
---

## Personal Information
Please fill out the following.
```{r}
name = "Sereen Kallerackal"
umid = "22465635"
```

## Load packages here.
```{r}
library(tidyverse)
library(haven)
library(NHANES)
library(ggplot2)
library(dplyr)
```


## Reading the "The 1999-2000 NHANES demographics data" file and "The 2015-2016 NHANES demographics data"
```{r}
data1 <- read_xpt(
  'NHANES 1999-2000 Demographics Data.xpt',
  col_select = NULL
)

data2 <- read_xpt(
  'NHANES 2015-2016 Demographics Data .xpt',
  col_select = NULL
)

```

## Calculating the total number of participants in "The 1999-2000 NHANES demographics data" and "The 2015-2016 NHANES demographics data"
```{r}
len1 <- print(length(data1$DMDHRAGE))
len2 <- print(length(data2$DMDHRAGE))

```

## Calculating summary of ages in "The 1999-2000 NHANES demographics data" and "The 2015-2016 NHANES demographics data"
```{r}
smry1 <- summary(data1$DMDHRAGE, na.rm = TRUE)
smry2 <- summary(data2$DMDHRAGE, na.rm = TRUE)

smry1

smry2

```

## Creating a Bar Plot of "The 1999-2000 NHANES demographics data" 
```{r}
barplot(table(data1$DMDHRAGE), main = "Age distribution", xlab="Age")

```

## Creating a Bar Plot of "The 2015-2016 NHANES demographics data 
```{r}
barplot(table(data2$DMDHRAGE), main = "Age distribution", xlab="Age")

```

## SUMMARY
The 1999-2000 NHANES demographics data has 9965 participants and the 2015-2016 NHANES demographics data has 9971 participants.
Additionally, the 1999-2000 bar plot shows a greater number of frequency between the ages 33 and 41. You can see a downward trend in this distribution. The 2015-2016 bar plot shows more frequency between the ages 45 and 78. And there is a sharp rise in age 80 above. 
This anlysis provides evidence to show an increase in the general population age. 












