---
title: "LHS 610: Week 1 Lab Assignment"
author: "Sereen Kallerackal"
date: "22 January 2021"
output: html_notebook
---

## Submission Instructions
Your lab solutions should be written entirely in this R notebook file. Once you are completed with the lab assignment, you should submit (1) the .nb.html file and (2) the .csv file generated at the very end through Canvas (lhs610_lab1.csv). Before submitting, please make sure to run all code chunks and that all outputs are visible in the .nb.html file. Also make sure that all the code chunks execute without any errors.

## Grading
We will be using a combination of autograding and manual grading to grade your lab assignment submissions. Questions requiring you to store your answer to a variable  and labeled as "AUTOGRADED" will be autograded, while more open-ended questions will be manually graded by looking through the code and outputs in your .nb.html file. 

## Week 1 Lab Assignment Instructions
Using the NHANES data frame from the NHANES package, answer the following questions. Each question should be answered using code rather than looking at RStudio’s user interface. In addition to your code, please provide an answer to the question if it is not readily apparent from your code’s output. The first two questions have been done for you as an example. 

## Personal Information
Please fill out the following.
```{r}
name = "Sereen Kallerackal"
umid = "22465635"
```

## Load any necessary packages here.
```{r}
library(tidyverse)
library(NHANES)
help(package= 'NHANES')
data("NHANES")
# Load any additional packages here
```


1. AUTOGRADED How many rows and columns are in the dataset? Please store the number of rows and columns to the variables `q1_row` and `q1_col` respectively. 
```{r}
q1_row = nrow(NHANES)
q1_row

q1_col = ncol(NHANES)
q1_col
```

2. AUTOGRADED What is the distribution (6-number summary) for the systolic blood pressure or SBP (first reading only)? Fill out the blanks for the variable `q2_summary`.
```{r}
summary(NHANES$BPSys1)

# Fill out the blanks here
q2_summary = "Min = 72
              1st Qu. = 106
              Median = 116
              Mean = 119.1
              3rd Qu. = 128
              Max = 232
              NA's = 1763"
```

3. AUTOGRADED Directly calculate the mean and median systolic blood pressure (first reading only) using the mean() and median() functions. Please store the mean and median to the variables `q3_mean` and `q3_median` respectively. 
* If you get a result that you didn’t expect, take a look at the documentation
```{r}
# YOUR CODE HERE
# Not sure why I am getting an NA
q3_mean = mean(NHANES$BPSys1, na.rm = TRUE)
q3_mean
q3_median = median(NHANES$BPSys1, na.rm = TRUE)
q3_median 
```

4. AUTOGRADED What is the 10th and 90th percentile for systolic blood pressure (first reading only)? Please store the results of the 10th and 90th percentile to the variables `q4_10` and `q4_90` respectively. 
* You can look online to find a function that will help you get this information
```{r}
# YOUR CODE HERE
q4_10 = quantile(NHANES$BPSys1,probs = 0.1, na.rm = TRUE)
q4_10
q4_90 = quantile(NHANES$BPSys1,probs = 0.9, na.rm = TRUE)
q4_90
  
```

5. AUTOGRADED What is the second highest systolic blood pressure (first reading only)? How many people have this SBP? Store the second highest SBP and number of people with the SBP to the variables `q5_sbp_high` and `q5_sbp_num` respectively. 
```{r}
# YOUR CODE HERE
table(NHANES$BPSys1)
q5_sbp_high = 224 
q5_sbp_num = 2 

```

6. AUTOGRADED What is the breakdown for education level for people in this dataset? How many people are in each type of group? Fill out the blanks for the variable `q6_edu_lev`.
```{r}
# YOUR CODE HERE
table(NHANES$Education)

# Fill out the blanks here
q6_edu_lev = "8th Grade = ___
              9-11th Grade = ___
              High School = ___
              Some College = ___
              College Grad = ___"
```

## Please run, but DO NOT modify this section to generate a table with your answers for submission. You can also use this table as a way to double check your answers. 
```{r}
lab1_answers = tibble(name, umid, q1_row, q1_col, 
                      q2_summary, 
                      q3_mean, q3_median, 
                      q4_10, q4_90,
                      q5_sbp_high, q5_sbp_num, 
                      q6_edu_lev)
write_csv(lab1_answers, "lhs610_labw1.csv")
```